---
title: "Analysis of Titanic survival"
author: "Ameyaw Ahmed Analysis"
date: '2022-07-14'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# HYPOTHETICAL SCENARIO
 You are a consulting data scientist and have been hired to analyse the titanic data.
 - The goal of this analysis is to explain patterns of survival in the data

 - The audience is decidedly non-data savvy!
 
```{r loading the packages}
library(tidyverse)
library(tidyquant)
library(scales)
```

```{r loading data}
titanic <- read_csv(here::here("raw_data", "titanic.csv"))
```

```{r viweing the data}
head(titanic, n = 15)
```

# Data wrangling
- Setting variables as factors
```{r}
titanic$Pclass <- as_factor(titanic$Pclass)
titanic$Survived <- as_factor(titanic$Survived)
titanic$Sex <- as_factor(titanic$Sex)
titanic$Embarked <- as_factor(titanic$Embarked)
```

## Questions to answer

 - 1. What was the survival rate

```{r}
ggplot(data = titanic, aes(x = Survived)) +
  geom_bar() +
  theme_bw() +
  labs(x = "survived", y = "passengers count") +
  ggtitle("Titanic survival rate")
```

- 2. What was the percentage rate by survival
```{r}
ggplot(data = titanic, aes(x =Sex, fill = Survived)) +
  geom_bar() +
  theme_bw() +
  labs(x = "gender", y = "passengers count") +
  theme(legend.position = "top") +
  ggtitle("Survival rate by gender")
```

- 3. What was the survival rate by class of ticket
```{r}
ggplot(data = titanic, aes(x =Pclass, fill = Survived)) +
  geom_bar() +
  theme_bw() +
  labs(x = "class", y = "passengers count") +
  theme(legend.position = "top") +
  ggtitle("Survival rate by ticket class")
```

- 4. What was the survival rate by class of ticket and gender
```{r}
ggplot(data = titanic, aes(x = Sex, fill = Survived)) +
  geom_bar() +
  facet_wrap(~ Pclass) +
  theme_bw() +
  labs(x = "gender", y = "passengers count") +
  theme(legend.position = "top") +
  ggtitle("Survival rate by gender and class")
```

- 5. What is the distribution of passenger age
```{r}
ggplot(data =  titanic, aes( x = Age )) +
  geom_histogram(binwidth = 5) +
  theme_bw() +
  labs(x = "age",
       y = "passenger count", 
       title = "Age distribution")
```

- 6. What are the survival rates by age?
```{r}
ggplot(data =  titanic, aes( x = Age, fill = Survived )) +
  geom_histogram(binwidth = 5) +
  theme_bw() +
  labs(x = "age",
       y = "passenger count", 
       title = "Age distribution")
```

```{r}
ggplot( data = titanic, aes(x = Survived, y = Age)) +
  geom_boxplot() + 
  theme_classic() +
  labs(x = "Survived", 
       y = "age", 
       title = "Distribution of age by survival")
```

-7. What is the survival rate by age when segmented by gender and class of ticket
```{r}
ggplot( data = titanic, aes(fill = Survived, x = Age)) +
  facet_wrap(Sex ~ Pclass) +
  geom_density(alpha = 0.5) + 
  theme_bw() +
  labs(x = "Survived", 
       y = "age", 
       title = "Survival rate by Age, sex and ticket class")
```


```{r}
prop.table(table(titanic$Survived))
```


